Dynamic parameter estimation in the nested multinomial logit choice model

### 2021-08-24 paper/paper.tex
rev

\begin{abstract}
This paper considers dynamic parameter estimation in the nested multinomial logit choice model under a tight computational constraint. We propose a scheme that updates the current estimate according to an intuitive step rule. Stochastic convergence is guaranteed under the Robbins--Monro condition. Numerical simulations demonstrate the technique's feasibility.
\end{abstract}

@@

This paper is structured as follows. Sections \ref{modeldescription} and \ref{staticestimation} establish the notation and preliminary results in choice modeling. Section \ref{dynamicestimation} describes the update step and shows that it is a Robbins--Monro algorithm. Section \ref{computation} provides numerical results.

@@

A known issue with the MNL choice model is the assumption that introducing a new item into the assortment does not alter the preferability of the items already in the assortment relative to one another. This assumption, known as the independence of irrelevant alternatives, is violated when some products are near substitutes. The \emph{nested} multinomial logit model (NMNL) addresses this issue by dividing the consumer’s decision into two stages.

@@

\begin{proposition}\label{mnlupdateconverges}If the observations are drawn from an MNL choice model and each consideration set appears with nonzero probability, then the MNL parameter update \eqref{mnlparameterupdate} converges in expectation to the true parameters $\delta$. 
\end{proposition}

@@

\begin{proposition}If the observations are drawn from an NMNL choice model and each consideration set appears with nonzero probability, then the NMNL parameter update converges to the true parameters $\sigma$ and $\delta$.\end{proposition}

\section{Computational examples} \label{computation}
We demonstrate the tractability of the dynamic parameter update in three simulations. The first simulation concerns the MNL case and is summarized in Figure \ref{param-est-dynamic-nonnested}, while the second concerns the NMNL case and is summarized in Figure \ref{param-est-dynamic-nested}. In both simulations, a small number of products was chosen to yield a legible graph; the parameter update formula is computationally tractable for much higher values of $n$ and $m$. The third simulation considers a larger product portfolio with NMNL choice, and it compares the accuracy of maximum likelihood estimation over a big dataset with the final estimate that results from treating the same observations as a data stream. While the accuracy is necessarily lower in the latter case, it may be acceptable given the savings in computation time. In all cases, it is possible to obtain smoother convergence by taking a rolling average or tuning the step parameters $\alpha$, $\beta$, and $r$. The Julia code for these simulations is available on Github at \url{https://github.com/maxkapur/MultinomialLogit}.

@@

%% else use the following coding to input the bibitems directly in the
%% TeX file.

>>>>

\begin{abstract}
This paper considers dynamic parameter estimation in the nested multinomial logit choice model under a tight computational constraint. We propose an intuitive step rule that updates the parameters each time a new consumer choice is observed. Stochastic convergence is guaranteed under the Robbins--Monro condition, and numerical simulations demonstrate the technique's feasibility.
\end{abstract}

@@

This paper is structured as follows. Sections \ref{modeldescription} and \ref{staticestimation} establish the notation and preliminary results in choice modeling. Section \ref{dynamicestimation} describes the update step and shows that it is a Robbins--Monro algorithm. Finally, section \ref{computation} provides numerical results.

@@

A known issue with the MNL choice model is the assumption that introducing a new item into the assortment does not affect the preferability of the other items in the assortment relative to one another. This assumption, known as the independence of irrelevant alternatives, is violated when some products are near substitutes. The \emph{nested} multinomial logit model (NMNL) addresses this issue by dividing the consumer’s decision into two stages.

@@

\begin{proposition}\label{mnlupdateconverges} If the observations are drawn from an MNL choice model and each consideration set appears with nonzero probability, then the MNL parameter update \eqref{mnlparameterupdate} converges in expectation to the true parameters $\delta$. 
\end{proposition}

@@

\begin{proposition} \label{nmnlupdateconverges}
If the observations are drawn from an NMNL choice model and each consideration set appears with nonzero probability, then the NMNL parameter update converges to the true parameters $\sigma$ and $\delta$.\end{proposition}

Propositions \ref{mnlupdateconverges} and \ref{nmnlupdateconverges} require that each consideration set appear with nonzero probability in order to ensure that the input data contains information about the relative preferability of all possible pairs of products. The full-support assumption may be relaxed to the requirement that the graph $\mathcal{G}$ is connected, where the nodes of $\mathcal{G}$ are the products $\mathcal{N}$, and $\mathcal{G}$ has an arc between $i$ and $j$ if and only if both products belong to a consideration set that appears with nonzero probability. Indeed, the numerical experiments below employ this weaker assumption.

\section{Computational examples} \label{computation}
We demonstrate the tractability of the dynamic parameter update in three simulations. The first simulation concerns the MNL case and is summarized in Figure \ref{param-est-dynamic-nonnested}, while the second concerns the NMNL case and is summarized in Figure \ref{param-est-dynamic-nested}. In both simulations, a small number of products was chosen to yield a legible graph; the parameter update formula is computationally tractable for much higher values of $n$ and $m$. The third simulation considers a larger product portfolio with NMNL choice, and it compares the accuracy of maximum likelihood estimation over a big dataset with the final estimate that results from treating the same observations as a data stream. While the accuracy is necessarily lower in the latter case, it may be acceptable given the savings in computation time. In all cases, it is possible to obtain smoother convergence by taking a rolling average or tuning the step parameters $\alpha$, $\beta$, and $r$.

The Julia code for these simulations is available on Github at \url{https://github.com/maxkapur/MultinomialLogit}.

@@

%% else use the following coding to input the bibitems directly in the
%% TeX file.ｑｓ
